Zero-Markov information in topological games

### 2015-11-16 articles/zero-markov-information-in-a-topological-game/article.tex
revisions per the referee

\begin{theorem}
  The following are equivalent for a locally compact space $X$:
    \begin{itemize}
      \item $X$ is paracompact
      \item $\pl K \win \gruKLGame{X}$. (\(\pl K\) has a winning strategy
      for the game.)
    \end{itemize}
\end{theorem}

@@

\begin{theorem}
  The following are equivalent for a locally compact space $X$:
    \begin{itemize}
      \item $X$ is metacompact
      \item $\pl K \tactwin \gruKPGame{X}$ (\(\pl K\) has a \term{tactical}
      winning strategy which only considers the most recent move of
      the opponent each round)
    \end{itemize}
\end{theorem}

@@

\begin{theorem}
  The following are equivalent for a locally compact space $X$:
    \begin{itemize}
      \item $X$ is $\sigma$-metacompact
      \item $\pl K \markwin \gruKPGame{X}$ (\(\pl K\) has a \term{Markov}
      winning strategy which only considers the most recent move of
      the opponent and the round number each round)
    \end{itemize}
\end{theorem}

@@

\begin{proof}
  Let \(K_n\) for \(n<\omega\) witness hemicompactness.
  If \(C\cap K_n\) is closed for each \(n<\omega\), then let \(K\) be any compact
  set. Since \(K\subseteq K_n\) for some \(n<\omega\), \(C\cap K\) is closed, and
  therefore \(C\) is closed.
\end{proof}

@@

@@

From these definitions it is easily verified that principal ultrafilters
are isolated, so \(\kappa\) with the discrete topology may be viewed as
a dense open subspace of \(\beta\kappa\).
We wish to consider the subspace
of \(\beta\omega\) consisting of all principal ultrafilters and a single free
ultrafilter \(\mc F\), denoted by \(\omega\cup\{\mc F\}\).

@@

\begin{proof}
  Let \(I=\{n_i:i<\omega\}\cup\{\mc F\}\) be infinite.
  Then \(\{U_{\omega\setminus\{n_i:i\geq j\}}:j<\omega\}\) is an open cover of
  \(I\cup\{\mc F\}\) with no finite subcover.
\end{proof}

@@

\begin{proof}
  Let \(\sigma\) be a predetermined strategy for \(\pl K\), and define the legal
  counter-attack \(H:\omega\to\mc K(X)\) by
  \(H(n)=(n\cup\sigma(n+1))\setminus\sigma(n)\). Then for any neighborhood
  \(U_S\) of \(\mc F\), \(S\) is infinite, and since
  \(\bigcup_{n<\omega} H(n)\supseteq\omega\setminus\sigma(0)\), \(U_S\) meets
  infinitely many of the finite \(H(n)\). Thus \(\sigma\) is not a winning
  predetermined strategy.
\end{proof}

@@

It is not possible to prove in \(ZFC\) that
\(\pl K\prewin \gruKPGame{\omega\cup\{\mc F\}}\)
for arbitrary free ultrafilters.

@@

\begin{theorem}
  CH implies the existence of a selective ultrafilter.
  \cite{MR0080902}
\end{theorem}

@@

Of particular note is that the author knows of no examples of a
non-\(k\)-space such that \(K\prewin\gruKPGame{X}\).

\begin{question}
  Does \(K\prewin\gruKPGame{X}\) imply \(X\) is a \(k\)-space?
\end{question}

>>>>

\begin{theorem} \cite[Theorem 5]{MR858337}
  The following are equivalent for a locally compact space $X$:
    \begin{itemize}
      \item $X$ is paracompact
      \item $\pl K \win \gruKLGame{X}$. (\(\pl K\) has a winning strategy
      for the game.)
    \end{itemize}
\end{theorem}

@@

\begin{theorem} \cite[Theorem 2]{MR858337}
  The following are equivalent for a locally compact space $X$:
    \begin{itemize}
      \item $X$ is metacompact
      \item $\pl K \tactwin \gruKPGame{X}$ (\(\pl K\) has a \term{tactical}
      winning strategy which only considers the most recent move of
      the opponent each round)
    \end{itemize}
\end{theorem}

@@

\begin{theorem} \cite[Theorem 3]{MR858337}
  The following are equivalent for a locally compact space $X$:
    \begin{itemize}
      \item $X$ is $\sigma$-metacompact
      \item $\pl K \markwin \gruKPGame{X}$ (\(\pl K\) has a \term{Markov}
      winning strategy which only considers the most recent move of
      the opponent and the round number each round)
    \end{itemize}
\end{theorem}

@@

\begin{proof}
  Let \(\{K_n:n<\omega\}\) be a collection of compact subsets
  which witnesses the hemicompactness of the space.
  Suppose \(C\cap K_n\) is closed for each \(n<\omega\).
  Let \(K\) be any compact subset of the space.
  There exists some \(n<\omega\) such that \(K\subseteq K_n\).
  Then \(C\cap K=(C\cap K_n)\cap K\) is closed. Therefore \(C\)
  is closed by the definition of \(k\)-space.
\end{proof}

@@

In particular note that for any ultrafilter on \(\kappa\) and subset
\(S\subseteq\kappa\), exactly one of \(S,\kappa\setminus S\) belongs to
the ultrafilter.

@@

From these definitions it is easily verified that principal ultrafilters
are isolated, so \(\kappa\) with the discrete topology may be viewed as
a dense open subspace of \(\beta\kappa\), with \(\beta\kappa\setminus\kappa\)
representing the free ultrafilters on \(\kappa\).
In that case, one might interpret \(U_S\) as
\(
  \{\mc F\in\beta\kappa\setminus\kappa:S\in\mc F\}
    \cup
  \{\alpha\in\kappa:\alpha\in S\}
\).

We wish to consider the subspace
of \(\beta\omega\) consisting of all principal ultrafilters and a single free
ultrafilter \(\mc F\), denoted by \(\omega\cup\{\mc F\}\).

@@

\begin{proof}
  Let \(n_i<\omega\) be distinct for each \(i<\omega\).
  Since \(\{n_i:i<\omega\}\) is infinite and discrete, it
  cannot be compact.
  So we consider \(I=\{n_i:i<\omega\}\cup\{\mc F\}\).
  Let \(S=\{n_{2i}:i<\omega\}\). If \(S\in \mc F\), then \(U_S\) is a
  neighborhood of \(\mc F\) missing \(\{n_{2i+1}:i<\omega\}\), so \(I\)
  contains a closed infinite discrete set and cannot be compact.
  Otherwise \(\omega\setminus S\in \mc F\) and \(U_{\omega\setminus S}\)
  is a neighborhood of \(\mc F\) missing \(\{n_{2i}:i<\omega\}\); again
  \(I\) contains a closed infinite discrete set and cannot be compact.
\end{proof}

@@

\begin{proof}
  Let \(\sigma\) be a predetermined strategy for \(\pl K\), and define the legal
  counter-attack \(H:\omega\to\mc K(X)\) by
  \(H(n)=(n\cup\sigma(n+1))\setminus\sigma(n)\). Then for any neighborhood
  \(U_S\) of \(\mc F\), \(S\) is infinite, and since
  \(\bigcup_{n<\omega} H(n)\supseteq\omega\setminus\sigma(0)\)
  and \(|H(n)|<\omega\), \(U_S\) meets
  infinitely many of the \(H(n)\). Thus \(\sigma\) is not a winning
  predetermined strategy.
\end{proof}

@@

We now show that it is not possible to prove in \(ZFC\) that
\(\pl K\prewin \gruKPGame{\omega\cup\{\mc F\}}\)
for every free ultrafilter; assuming the continuum hypothesis (\(CH\)),
we may find a counter-example.

@@

\begin{theorem}
  \cite{MR0080902}
  CH implies the existence of a selective ultrafilter.
\end{theorem}

@@

Of particular note is that the author knows of no examples of a
non-\(k\)-space such that \(\pl K\prewin\gruKLGame{X}\).

\begin{question}
  Does \(\pl K\prewin\gruKLGame{X}\) imply \(X\) is a \(k\)-space?
\end{question}
